%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Numerical Methods for CSE, ETH Zuerich
%
% (C) Seminar for Applied Mathematics, D-MATH
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{report}
% Specific information for course
\newcommand{\lectyear}{2015}
\newcommand{\lectterm}{AT'15}
\newcommand{\lecturer}{Prof. Ralf Hiptmair}
\newcommand{\lecture}{NumCSE}

%\providecommand{\slidemode}{4}
\usepackage{samlecturenotes}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Version with Python programs (created by F. Thaler, AS 2015)
% MATLAB programs ported to Python and included in lecture document.
% Inclusion of python specific parts controlled by boolean switch
\newboolean{withpython}
\setboolean{withpython}{true}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


% Graphics input path
\graphicspath{
{./Images/}
{./PICS/}
{./p1_SystemsOfEquations/Triangulations/}
{./p1_SystemsOfEquations/ch1_MatVec/PICTURES/}
{./p1_SystemsOfEquations/ch2_DirectMethodsLSE/PICTURES/}
{./p1_SystemsOfEquations/ch3_IterativeMethodsNonLinear/PICTURES/}
{./p1_SystemsOfEquations/ch4_IterativeMethodsLSE/PICTURES/}
{./p1_SystemsOfEquations/ch5_Eigenvalues/PICTURES/}
{./p1_SystemsOfEquations/ch6_LeastSquares/PICTURES/}
{./p1_SystemsOfEquations/ch7_Convolution/PICTURES/}
{./p1_SystemsOfEquations/ch8_FFT/PICTURES/}
{./p2_InterpolationApproximation/ch1_PolynomialInterpolation/PICTURES/}
{./p2_InterpolationApproximation/ch2_PiecewisePolynomials/PICTURES/}
{./p2_InterpolationApproximation/ch4_NumericalQuadrature/PICTURES/}
{./p2_InterpolationApproximation/ch6_ClusteringTechniques/PICTURES/}
{./p2_InterpolationApproximation/ch3_TrigonometricInterpolation/PICTURES/}
{./p2_InterpolationApproximation/ch5_FilteringWavelets/PICTURES/}
{./p3_ODE/ch1_SingleStepMethods/PICTURES/}
{./p3_ODE/ch2_ConvergenceStability/PICTURES/}
{./Maturandentag/PICTURES/}
}

\externaldocument{./NCSENEW}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Include macros
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{NCSENEW_macros}

% Partial compilation when in editing mode (selective inclusion)
%\ifthenelse{\equal{4}{\slidemode}}
%{%
%  \includeonly{NCSENEW_Testproblem}{}
%}{}

% Import SVN revision

\input{rev}

\begin{document}

\begin{tcolorbox}[colframe=black]
\begin{minipage}[t][0.6\textwidth][c]{\textwidth}
  ETH Lecture 401-0663-00L  Numerical Methods for CSE 
  \B\vspace{1.5cm}
  
  \begin{center}
    \Red{\Huge Homework Problems}
  \end{center}
  \vspace{1.5cm}
  
  \begin{center}
  \begin{large}
    {\Large Prof. R. Hiptmair, SAM, ETH Zurich}\\[1.5ex]

    Autumn Term 2016 %  draft version \svnToday, SVN rev. \svnInfoRevision

    (C) Seminar f\"ur Angewandte Mathematik, ETH Z\"urich
    \samskip

    URL: \href{http://www.sam.math.ethz.ch/~hiptmair/tmp/NumCSE/NumCSE15.pdf}{%
      http://www.sam.math.ethz.ch/\symbol{126}hiptmair/tmp/NumCSE/NumCSE15.pdf}
  \end{large}
\end{center}
\end{minipage}
\end{tcolorbox}
\bigskip

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% General information
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\fboxsep1ex
\fbox{\parbox{0.95\textwidth}{
    A steady and persistent effort spent on homework problems is essential for
    success in the course.
}}
\bigskip

You should expect to spend 4-6 hours per week on trying to solve the homework
problems. Since many involve small coding projects, the time it will take an
individual student to arrive at a solution is hard to predict.  
\bigskip

\begin{itemize}
\item The assignment sheets will be uploaded on the course
  \href{http://www2.math.ethz.ch/education/bachelor/lectures/hs2015/math/nummath_cse/index}{webpage}
  on Thursday every week.
\item Some or all of the problems of an assignment sheet will be discussed in the
  tutorial classes on Monday 1$\frac{1}{2}$ weeks after the problem sheet has been
  published.
\item A few problems on each sheet will be marked as {\bf core problems}. 
  Every participant of the course is strongly advised to try and solve \emph{at least}
  the core problems. 
\item If you want your tutor to examine your solution of the current problem
  sheet, please put it into the plexiglass trays in front of HG G 53/54 by the
  Thursday after the publication. You should submit your codes using the
  \href{https://people.math.ethz.ch/~grsam/submit/}{online submission
    interface}. This is voluntary, but feedback on your performance on homework
  problems can be important.
\item You are encouraged to hand-in incomplete and wrong solutions, since 
  you can receive valuable feedback even on incomplete attempts.
\item Please clearly mark the homework problems that you want your tutor to
  examine.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Selective inclusion of Problems
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\include{./Chapters/MatrixVector/gramschmidteigen}

\include{./Chapters/MatrixVector/ArrowMatrixVector}

\include{NCSENEW_Testproblem}%

\section{Inclusion of Codes}

Parts of a C++ code can be included by means of the \verb|\lstincludecpp| command:

Included with tag 2:

\begin{samcode}[C++ code]{cpp:1}{A demo code}
  \lstincludecpp[emph={NPDE},linebackgroundcolor={\lstcolorrange{2}{4}}]{test.cpp}{2}
\end{samcode}

Note \cref{l:1}! Included with tag 4, same file as \cref{cpp:1}.

\lstincludecpp[emph={Constant},linebackgroundcolor={\lstcolorlines{1,3,5}}]{test.cpp}{4}

\chapter{Summary}
\Label{cha:sum}

\cref{prb:test}: This is a test problem for demonstration only


\bibliographystyle{plain} 
\bibliography{lit}

\end{document}

